---
title: "Preprocessing Templae"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

## Information

This is the pre-processing template for the Emma Neuroscience Group. This template provides examples on how you can organize pre-processing in your project using R with RStudio.

RStudio is a wrapper around R, providing you with a visualized program (GUI: graphical user interface) that you can use to work easier with R. 

One feature of RStudio is the use of RMarkdown (.Rmd files). 

RMarkdown allows you to combine into one document:
- background information that will not be interpreted as part of the script (this text)
- the script text itself (code)
- the results produced by the code (e.g. figures)

The above can be integrated into one output document (html or pdf). 

This is an RMarkdown document, you are currently reading the background text that is not intepreted as code by R. You can run
- the script completely (Run > Run all) 
- parts of the script in isolation, organized in chunks starting with (play button)
- single lines using CTRL + ENTER


```{r setting, include=FALSE}

# setting for RMarkdown
knitr::opts_chunk$set(echo = TRUE)


# this line is necessary for working in the Azure environment
options(download.file.method="libcurl", repos=c("CRAN"="https://cran.r-project.org/"))

```

## Loading packages

At the beginning of each script, you will be loading the packages that you need. Packages are a set of scripts (functions) that is contributed open source by a fellow-user. Loading a package means that you add it to your session so that you can access the functions in the package.

We will load the packages using a function that we will call ipak. A function is a part of script that you can re-use throughout the remainder of the script once you have defined it. You could say it works as a hyperlink to a part of script that you may want to use repeatedly. It is worthwile to define functions yourself for operations that you want to repeat within a manuscript: if you want to change something in the function, you will only have to do it once.

Now let's define the function and load a package (dplyr and ggplot). If the package is not installed, then the function will first install the package and then load it.

```{r packages, include=FALSE}

# ipak function: install and load multiple R packages.
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}


# make a vector of strings with the packages that you want to load
packages <- c("dplyr", "ggplot2")

# feed the vector to the function ipak and the packages will be installed and loaded
ipak(packages)

```


## Source scripts

We will now start to source other Rmd scripts. This allows you to keep a better overview of your operations. 
- Please open the relevant script to find further information about the procedures performed.
- Find the results in the 'results' folder 


```{r create_database, echo=FALSE, include=FALSE}

### Merge ###
rmarkdown::render("scripts/Merge.Rmd", output_file = "../results/1. Merge.html")
#############



### Explore ###
rmarkdown::render("scripts/Explore.Rmd", output_file = "../results/2. Explore.html")
###############



### Outliers ###
# Define p-value for outliers
p <- .025  

# define the variables that you want to inspect
DV_inspect <- colnames(select(data, -c("subj", "age", "sex_male", "edu")))  

# Inspect
rmarkdown::render("scripts/Outliers.Rmd", output_file = "../results/3. Outliers.html")

# define the variables that you want to correct
DV_correct <- c("mrt_neutral", "exec_acc")  

# Correct outliers
rmarkdown::render("scripts/Winsorize.Rmd", output_file = "../results/3. Winsorize.html")
################



### Normality ###
data <- data_win  # Select input file
rmarkdown::render("scripts/Normality.Rmd", output_file = "../results/4. Normality.html")
#################



### Normalize ###
DV_norm <- c("d_proc", "mrt_neutral", "sd_neutral", "exec_acc")  # set the variables for normalization
rmarkdown::render("scripts/Normalize.Rmd", output_file = "../results/5. Normalize.html")
#################



### Missing values  ###
data <- data_norm # select input file
rmarkdown::render("scripts/Missing.Rmd", output_file = "../results/6. Missing.html")
# add missing per patient
#######################


### Impute  ###
data <- data_norm # select input file
omit <- c("subj") # select the variables to omit

rmarkdown::render("scripts/Impute.Rmd", output_file = "../results/7. Impute.html")
################



### PCA ###
f <- 9  # number of factors to extract
data <- data_imp_agg # select input file for the pca model
omit <- c("subj", "d_proc", "d_vis_mem", "d_ver_mem", "d_vis_wm", "d_ver_wm", "d_visuom", "age", "sex_male", "edu", "ver_mem_ret", "vis_mem_ret", "SCAT_tot")
rmarkdown::render("scripts/PCA.Rmd", output_file = "../results/8. PCA.html")
############



### Apply PCA ###
data <- data_imp_agg # select input file
rmarkdown::render("scripts/Apply_PCA.Rmd", output_file = "../results/9. Apply PCA.html")
##################



```
